---
title: "vignette_mtb_markdown"
author: "Y. Hsu"
date: '`r Sys.Date()`'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette_mtb_rmd}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mtb)
```

## Background

R Markdown has been popular in many communities for presenting and sharing results. 
There formatting functions help to emphasis a sentence, or a paragraph.
`add_colored_str()` adds a inline string with customized styles, including, color, background color and size.
`add_colored_box()` adds a colored area with title to emphasis a reminder, or a warning.

## How to use

This is a basic example which shows you how to format a string:

```{r example_rmd_1}
add_colored_str('Example - Warning', color=c(255,0,0), bold=TRUE, fontsize=1.5)
```

This is an example which shows you how to add a default box:

```{r example_rmd_2}
add_colored_box(info='additional information')
```

This is an example which shows you how to add a reminder box:

```{r example_rmd_3}
add_colored_box(type='green-reminder', info='Here is the reminding message.', width=0.7)
```

This is an example which shows you how to add a warning box:

```{r example_rmd_4}
add_colored_box(type='yellow-warning', info='Here is the warning message.', width=0.9)
```

This is an example which shows you how to add a warning box with stop sign and align to the right:

```{r example_rmd_5}
add_colored_box(type='red-stop', info='Here is the warning message.', halign='r', width=0.2)
```
